---
title: "Untitled"
format: docx
editor: source
---

```{r,include=FALSE}
library(kableExtra)
library(tidyverse)

# Loading indicator data
all_data <- load("all_data.RData")
```

OpenPrescribing provides NHS prescribing data aggregated by month, and some off-the-shelf indicators. As a proxy for asthma prevalence, we have extracted Short Acting Beta Agonist (SABA) inhalers. @tbl-fields lists the fields reported for each indicator and their availability.

```{r}
#| echo: false
#| label: tbl-fields
#| tbl-cap: OpenPrescribing indicators and availability by data field.

count_NAs <- function(df) {
    sapply(df, function(x) sum(is.na(x)))
  }

df <- all_saba_raw |> 
  drop_na(denominator) 

a <- (1-count_NAs(df)/nrow(df)) |>
    enframe(name = "field",
            value = "available") |>
    head(24) |> 
    mutate(available = sprintf("%.0f%%", available * 100))

a |> kable()
```

Similarly, data for a specific drugs prescribed in any medical practice contain the following data fields:

```{r}
#| echo: false
#| eval: false
#| label: tbl-fields-presc
#| tbl-cap: OpenPrescribing indicators and availability by data field.

drug_data <- read_csv("spending-by-practice-0212000AA.csv")

df <- drug_data

a <- (1-count_NAs(df)/nrow(df)) |>
    enframe(name = "field",
            value = "available") |>
    head(24) |> 
    mutate(available = sprintf("%.0f%%", available * 100))

a |> kable()

```

Prescription data is available at the following aggregation levels from January 2020 to December 2024:

1.  Regional teams

2.  ICB

3.  Sub-ICB

4.  PCN

5.  GP Practice
